Moment Constraints Make Learning Latent-Variable Models Easier

### 2014-01-28 doc/icml2014/abstract.tex
Updated background

\begin{abstract} 
\end{abstract}

>>>>

\begin{abstract}

Parameter estimation for latent variable models is typically
  performed by optimizing the log-likelihood, which is prone to local
  optima due to its non-convexity. 
In this paper, we extend recent advances in the method of moments 
  to provide consistent parameter estimates for a general family of
  directed and undirected graphical models, including grid models
  (\figureref{approach}) and latent tree models.
The key idea in our approach is that, given the observation potentials
  between observed and latent variables, the marginal distribution of
  every clique is recoverable from the corresponding partial moments.
We characterize the polynomial sample complexity for models in this
  family.
For models outside our family, we show how our approach can constrain
  parameter search by providing partial information via posterior
  regularization. 
Finally, we provide empirical intuition via experiments.

\end{abstract}

### 2014-01-28 doc/icml2014/background.tex
Updated background

Let $\sG$ be a discrete graphical model with variables $V$, of which
  $O \subseteq V$ be observed and $H = V \setminus O$ be hidden.
Let $\Pa(x) \subset V$ be the parents of a variable $x$ in a directed
  model, and $\sN(x) \subset V$ be the neighbours of $x$ in an undirected
  model.
We will use $o_1, o_2, \cdots \in O$ to denote observed variables and
  $h_1, h_2, \cdots \in H$ to denote hidden variables.
For ease of exposition, we assume that the domain of all the observed
  variables is $[D]$ and the domain of all the hidden variables is
  $[K]$.

Define $\otimes$ to be the tensor product. In matrix representation, it
  is the Kronecker product.\expand

@@

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.

@@

\citet{anandkumar12moments} present an algorithm to solve three-view mixture models.

Assume means $M_1, M_2, M_3$. Observe that the moments of the model are $M_{123}$.

Applying a whitening transformation and get a orthogonal tensor decomposition.

Solved using the robust tensor power method \citet{anandkumar13tensor}.

In general, we call hidden variables with this property ``bottlenecks'',
  and define them as follows.
\begin{definition}(Bottleneck)
  A hidden variable $h$ is said to be a bottleneck if there exists at
  least three conditionally independent observed variables $o_1, o_2,
  o_3$. Let $\sB(h)$ be the set of conditionally independent observed
  variables for $h$.
\end{definition}

\begin{assumption}
  We assume that $M_1, M_2, M_3$ are each full rank, and that $\pi \succ
  0$\reword.
\end{assumption}

>>>>

Let $\sG$ be a discrete graphical model with variables $V$, of which
  $O \subseteq V$ be observed and $H = V \setminus O$ be hidden.
Let $\Pa(x) \subset V$ be the parents of a variable $x$ in a directed
  model, and $\sN(x) \subset V$ be the neighbours of $x$ in an undirected
  model.
We will use $o_1, o_2, \cdots \in O$ to denote observed variables and
  $h_1, h_2, \cdots \in H$ to denote hidden variables.
For ease of exposition, we assume that the domain of all the observed
  variables is $[d]$ and the domain of all the hidden variables is
  $[k]$.

Define $\otimes$ to be the tensor product.
To denote the moments of observed variables, we will use the notation
\begin{align*}
  M_i &\eqdef \Pr(x_i), &
  M_{ij} &\eqdef \Pr(x_i, x_j), &
  M_{ijk} &\eqdef \Pr(x_i, x_j, x_k),
\end{align*}
and so on.
We will view $M_i, M_{ij}, M_{ijk}, \cdots$ as tensors of
  order 1, 2, 3, \ldots in $\Re^D, \Re^{D\times D}, \Re^{D \times
  D \times D}$.
For the moments of hidden variables, we will use analogous notation
\begin{align*}
  P_i &\eqdef \Pr(h_i), &
  P_{ij} &\eqdef \Pr(h_i, h_j), &
  P_{ijk} &\eqdef \Pr(h_i, h_j, h_k).
\end{align*}

@@

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $\sD = (x^{(1)}, \dots, x^{(n)})$
where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.

@@

\begin{figure}[t]
  \label{fig:three-view}
  \centering
  \input{figures/three-view.tikz}
  \caption{A three-view mixture model}
\end{figure}

Recently, \citet{anandkumar12moments} presented a consistent algorithm
  to learn the parameters of a three-view mixture model
  (\figureref{three-view}) that sidesteps the local optima problem
  altogether. 
In the model, we have three observed variables $x_1, x_2, x_3 \in [d]$
  that are independent conditioned on the latent variable $h \in [k]$. 
  The parameters of this model are the conditional means $O^{(i)} \eqdef
  \E[x_i | h] \in \Re^{D \times K}$ for $i \in \{1,2,3\}$ and the mixture probabilities $\pi
  \in \Delta_{k-1}$.
  The key idea is that the observed moments $M_{12}, M_{23}, M_{13}$ and
  $M_{123}$ admit a factorization in terms of parameters;
\begin{align*}
  M_{ij} &= \sum_h \pi_h O^{(i)}_h \otimes O^{(j)}_h & \forall i,j \in \{1,2,3\} \\
  M_{123} &= \sum_h \pi_h O^{(1)}_h \otimes O^{(2)}_h \otimes O^{(3)}_h.
\end{align*}

\citet{anandkumar13tensor} presented a robust tensor power method, which
  we call $\TensorFactorize$, which converts consistent estimates of the observed moments 
  into consistent estimates of the parameters,
  $\pi, M_1, M_2, M_3$ (up to permutation of the columns), provided
  the following assumption holds;

\begin{assumption}
Assume the conditional moments $M_1, M_2, M_3$ have full column rank
  $k$, and that $\pi \succ 0$.
\end{assumption}

In brief, the algorithm first computes a whitening matrix $W \in
  \Re^{d \times k}$ such that $W^T M_{12} W = I_{k \times k}$,
  and uses it to transform $M_{123}$ into a symmetric orthogonal tensor.  
Then a robust tensor power method is applied to extract the eigenvectors
  of the whitened $M_{123}$; unwhitening yields the columns of $M_{3}$.
$M_1$ and $M_2$ can then be recovered with some algebra.

\todo{Mention {\em method of moments} somewhere?}
